---
title: "Examining the Influence of Urban Social Disorder Events on Deadly Incidents Across Geographic Regions"
author: "Emory Bayer"
institute: "American University"
date: "03/22/2024 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: 
      - "default"
      - "default-fonts"
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    class: 
      - "themer-better-default" # Choose a professional theme
    dev: 'svg'  # Specify SVG as the graphics device
---               

```{r setup, include = FALSE} 
options(htmltools.dir.version = FALSE)
knitr :: opts_chunk$set(
  echo = FALSE,
  message = FALSE, 
  warning = FALSE ,
  fig.align = 'center')

library(tidyverse)
library(knitr)
library(cowplot)
library(patchwork)
library(ggplot2)
library(readxl)
library(dplyr)


setwd("C:/Users/emory/Documents/Statistics Level 2")
df = read_excel("Copy of cities.xlsx")
```
Class: inverse, bottom  

#Topic

###Data:
####This dataset, from the Urban Social Disorder Global Dataset, analyzes the impact of urbanization on global populations. It covers 186 national capitals and urban centers from 1960 to 2014, detailing social disorder events, both violent and non-violent, including fatalities. The data is categorized by geographic region, longitude, and latitude. 

###Reason this dataset is important: 
####As the global population increasingly centralizes in cities, it becomes crucial to comprehend social unrest trends within these urban areas. For instance, understanding whether certain regions are more prone to experiencing violent social disorder outbreaks is essential for implementing policies aimed at mitigation.

---
# Analysis 1.
<small>This visual analysis examines social disorder events in the Middle East and North Africa, focusing on the relationship between the variables NEVENTS and DEATHEVENTS.<small>

**NEVENTS** = Total number of social disorder events occurring in the city between 1960 and 2014 

**DEATHEVENTS** = Total number of events resulting in at least one death occurring in the city between 1960 and 2014  

```{r, fig.width=7, fig.height=5}
Middle_E_NAfrica <- df %>% 
  filter(REGION == "Middle East and North Africa")

plot1 <- ggplot(Middle_E_NAfrica, aes(x = NEVENTS, y = DEATHEVENTS)) +
  geom_line(color = "purple") +
  geom_point() +
  scale_y_continuous(breaks = seq(0, max(Middle_E_NAfrica$DEATHEVENTS), by = 50)) +
  scale_x_continuous(breaks = seq(0, max(Middle_E_NAfrica$NEVENTS), by = 50)) +  
  labs(x = "Number of events", y = "Number of Deadly events", title = "Impact of Number of Social Disorder events on the number 
       of Deadly events in Middle East and North Africa") +
  theme_minimal(base_size = 9) +
  theme(axis.title.x = element_text(size = 13),
        axis.title.y = element_text(size = 13))  +  
  theme(
    plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),  
    plot.background = element_rect(fill = "white"),  
    panel.grid.major = element_line(color = "gray"),   
    panel.grid.minor = element_blank(),  
    axis.line = element_line(color = "black"), 
    axis.text = element_text(color = "black", size = 14), 
    axis.title = element_text(color = "black"),  
    plot.title = element_text(size = 18, hjust = 0.5),  
    axis.ticks = element_line(color = "black", size = 0.5)  
  )

print(plot1)
```



---
# Analysis 2

  **Impact of number of events on the number of **
  **deadly events in North Africa and Latin America**
```{r, fig.width=8, fig.height=7}
Middle_E_NAfrica <- df %>% 
  filter(REGION == "Middle East and North Africa")

plot1 <- ggplot(Middle_E_NAfrica, aes(x = NEVENTS, y = DEATHEVENTS)) +
  geom_line(color = "purple") +
  geom_point() +
  scale_y_continuous(breaks = seq(0, max(Middle_E_NAfrica$DEATHEVENTS), by = 50)) +
  scale_x_continuous(breaks = seq(0, max(Middle_E_NAfrica$NEVENTS), by = 100)) +
  labs(x = "Number of events", y = "Number of Deadly events") +  
  annotate("text", x = Inf, y = Inf, label = "North Africa and Middle East", size = 6, hjust = 1, vjust = 1) +
  theme(axis.title.y = element_text(size = 12),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(margin = margin(t = 15)))  

Latin_America <- df %>%
  filter(REGION == "Latin America")

plot4 <- ggplot(Latin_America, aes(x = NEVENTS, y = DEATHEVENTS)) +
  geom_line(color = "red") +
  geom_point() +
  scale_y_continuous(breaks = NULL, limits = c(0, 350), minor_breaks = seq(0, 350, by = 50)) +
  scale_x_continuous(breaks = seq(0, max(Latin_America$NEVENTS), by = 50)) +
  labs(x = "Number of events", y = NULL) +  
  theme(panel.grid.major = element_line(color = "white", linetype = "solid"),
        panel.grid.minor = element_line(color = "white", linetype = "solid"),
        axis.text = element_text(size = 12),
        axis.title.x = element_text(margin = margin(t = 15)), 
        axis.title.y = element_blank()) + 
  annotate("text", x = Inf, y = Inf, label = "Latin America", size = 6, hjust = 1.5, vjust = 1)

# Combine plots side by side using patchwork
combined_plots <- plot1 + plot4 +
  plot_layout(ncol = 2) +  
  theme(
    plot.margin = margin(40, 40, 40, 40),  
    axis.title = element_text(margin = margin(t = 20))
  )

print(combined_plots)
```

